---
title: "Derived FLUXNET variables"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(ingestr)
```

## Daytime VPD

Daytime VPD is not provided in the daily FLUXNET2015 data files. Instead, we can derive it based on the half-hourly data, averaging half-hourly VPD of time steps where the shortwave incoming radiation `"SW_IN_F"` is greater than zero (i.e., during the day). To do this, we can set the `getvars` argment to `list(vpd  = "VPD_F_DAY")` and additionally provide the directory where half-hourly files are located locally through the argument `settings`. We do this just for one year here (only year 2000 data is provided as external data in this package).
```{r message=FALSE}
settings_fluxnet <- list(dir_hh = paste0(path.package("ingestr"), "/extdata/"), getswc = FALSE)

df_fluxnet <- ingest_bysite(
  sitename = "FR-Pue",
  source = "fluxnet",
  getvars = list(vpd  = "VPD_F_DAY"),
  dir = paste0(path.package("ingestr"), "/extdata/"),
  settings = settings_fluxnet,
  timescale = "d",
  year_start = 2007,
  year_end = 2007,
  verbose = FALSE
  )
df_fluxnet
```

## Daily minimum temperature

Daily minimum temperature is not provided in the daily FLUXNET2015 data files. Instead, we can derive it based on the half-hourly data. To do this, we can set the `getvars` argument to `list(tmin  = "TMIN_F")` and additionally provide the directory where half-hourly files are located locally through the argument `settings`. We do this just for one year here (only year 2000 data is provided as external data in this package).
```{r message=FALSE}
settings_fluxnet <- list(dir_hh = paste0(path.package("ingestr"), "/extdata/"), getswc = FALSE)

df_fluxnet <- ingest_bysite(
  sitename = "FR-Pue",
  source = "fluxnet",
  getvars = list(tmin  = "TMIN_F"),
  dir = paste0(path.package("ingestr"), "/extdata/"),
  settings = settings_fluxnet,
  timescale = "d",
  year_start = 2007,
  year_end = 2007,
  verbose = FALSE
  )
df_fluxnet
```