---
title: "Examples for ingestr"
output: html_document
---

```{r setup, include=FALSE}
library(ingestr)
library(readr)
library(dplyr)
library(lubridate)
library(rsofun)
```

## Example for FLUXNET 2015 sites

Get site meta information and define a list of sites.
```{r}
mysites <- c("BE-Vie", "DE-Tha", "DK-Sor", "FI-Hyy", "IT-Col", "NL-Loo", "US-MMS", "US-WCr", "US-UMB", "US-Syv", "DE-Hai", "IT-MBo", "US-GLE", "FR-Fon", "NL-Hor", "US-UMd", "AU-Dry", "DE-Obe", "IT-Tor", "US-Wi4")

siteinfo <- read_csv("~/data/FLUXNET-2015_Tier1/siteinfo_fluxnet2015_sofun+whc.csv") %>% 
  rename(sitename = mysitename) %>% 
  filter(sitename %in% mysites) %>% 
  mutate(date_start = lubridate::ymd(paste0(year_start, "-01-01"))) %>% 
  mutate(date_end = lubridate::ymd(paste0(year_end, "-12-31")))
```

### rsofun input data

First read all FLUXNET meteo data.
```{r message=FALSE, warning=FALSE}
ddf_fluxnet <- ingest(
  siteinfo = siteinfo,
  source    = "fluxnet2015", 
  getvars   = list(temp = "TA_F_DAY", prec = "P_F", vpd  = "VPD_F_DAY", swin =  "SW_IN_F", netrad = "NETRAD", patm = "PA_F"),
  dir       = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
  settings  = list(dir_hh = "~/data/FLUXNET-2015_Tier1/20191024/HH/", getswc = FALSE, threshold_GPP = 0.5),
  timescale = "d"
  )
```

Some meteo data is not available from FLUXNET. Extract it from WATCH-WFDEI global climate reanalysis files instead.
```{r}
ddf_watch <- ingest(
  siteinfo = siteinfo,
  source    = "watch_wfdei", 
  getvars   = list(temp = "Tair"),
  dir       = "~/data/watch_wfdei/"
  )
```

Some meteo data is not available from FLUXNET. Extract it from CRU global climate files instead.
```{r}
ddf_cru <- ingest(
  siteinfo = siteinfo,
  source    = "cru", 
  getvars   = list(ccov = "cld"),
  dir       = "~/data/cru/ts_4.01/"
  )
```


### Evaluation data

```{r message=FALSE, warning=FALSE}
## get data for idfferent time scales separately
ddf_eval <- ingest(
  siteinfo = siteinfo,
  source    = "fluxnet2015", 
  getvars   = list(latenth = "LE_F_MDS", latenth_qc = "LE_F_MDS_QC"),
  dir       = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
  settings  = list(threshold_LE = 0.8, getswc = TRUE),
  timescale = "d"
  )
mdf_eval <- ingest(
  siteinfo = siteinfo,
  source    = "fluxnet2015", 
  getvars   = list(latenth = "LE_F_MDS", latenth_qc = "LE_F_MDS_QC"),
  dir       = "~/data/FLUXNET-2015_Tier1/20191024/MM/",
  settings  = list(threshold_LE = 0.8, getswc = TRUE),
  timescale = "m"
  )
adf_eval <- ingest(
  siteinfo = siteinfo,
  source    = "fluxnet2015", 
  getvars   = list(latenth = "LE_F_MDS", latenth_qc = "LE_F_MDS_QC"),
  dir       = "~/data/FLUXNET-2015_Tier1/20191024/YY/",
  settings  = list(threshold_LE = 0.8, getswc = TRUE),
  timescale = "y"
  )
```

Use rsofun to create a standardised object used for benchmarking the model output.
```{r}
settings_eval <- list(
  benchmark = list( latenth = c("fluxnet2015") )
  )
obs_eval <- get_obs_eval2( settings_eval = settings_eval, adf = adf_eval, mdf = mdf_eval, ddf = ddf_eval )
```

### Calibration data

tbc
