---
title: "Getting MODIS EVI data from Google Earth Engine"
author: "Koen Hufkens"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting MODIS EVI data from Google Earth Engine}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
library(ingestr)
library(ggplot2)
# library(readr)
# library(dplyr)
# library(lubridate)
# library(rsofun)
```


## Single site

### Prepare access to Google Earth Engine


The library `gee_subset` by Koen Hufkens can be downloaded from this [link](https://khufkens.github.io/gee_subset/) and used to extract data directly from Google Earth Engine. Note that this requires the following programmes to be available:

- git: You can use [Homebrew](https://brew.sh/) to installing git by entering in your terminal: `brew install git`.
- [python](https://www.python.org/)

Then, carry out the follwing steps:

- In your terminal, change to where you want to have the repository. In this example, we're cloning it into our home directory:
```{sh, eval = FALSE}
cd ~
git clone https://github.com/khufkens/google_earth_engine_subsets.git
```

To get access to using the Google Earth Engine API (required to use the `gee_subset` library), carry out the following steps in your terminal. This follows steps described [here](https://github.com/google/earthengine-api/issues/27).

1. Install google API Python client
```{sh, eval = FALSE}
sudo pip install --upgrade google-api-python-client
```
I had an error and first had to do this here following [this link](https://github.com/pypa/pip/issues/3165):
```{sh, eval = FALSE}
sudo pip install --ignore-installed six
```

2. Install pyCrypto
```{sh, eval = FALSE}
sudo pip install pyCrypto --upgrade
```

3. Install Python GEE API
```{sh, eval = FALSE}
sudo pip install earthengine-api
```

4. Run authentification for GEE
```{sh, eval = FALSE}
earthengine authenticate
```

5. Finally, try if it works. This shouldn't return an error:
```{sh, eval = FALSE}
python -c "import ee; ee.Initialize()"
```


### Download data

To facilitate the selection of data products and bands to be downloaded, you may use the function `get_settings_gee()` which defines defaults for different data bundles (`c("modis_fpar", "modis_evi", "modis_lai", "modis_gpp")` are available). We use `"modis_evi"`, downloading the MODIS/006/MOD13Q1, band EVI data.

The following example is for downloading MODIS EVI data.
```{r eval = FALSE}
settings_gee <- get_settings_gee(
  bundle            = "modis_evi",
  python_path       = system("which python", intern = TRUE),
  gee_path          = "~/google_earth_engine_subsets/gee_subset/",
  data_path         = "~/data/gee_subsets/",
  method_interpol   = "linear",
  keep              = TRUE,
  overwrite_raw     = FALSE,
  overwrite_interpol= TRUE
  )
```

This can now be used to download the data to the directory specified by argument `data_path` of function `get_settings_gee()`.
```{r eval = FALSE}
df_gee_modis_fpar <- ingest_bysite(
  sitename  = "CH-Lae",
  source    = "gee",
  year_start= 2010,
  year_end  = 2012,
  lon       = 8.365,
  lat       = 47.4781,
  settings  = settings_gee,
  verbose   = FALSE
  )
```

Plot this data.
```{r eval = FALSE}
plot_fapar_ingestr_bysite(df_gee_modis_fpar, settings_gee)
```


## Multiple sites

Using the same settings as specified above, we can download MODIS FPAR data for multiple sites at once from GEE:
```{r warning=FALSE, message=FALSE, echo = T, results = 'hide', eval = FALSE}
settings_gee <- get_settings_gee(
  bundle            = "modis_evi",
  python_path       = system("which python", intern = TRUE),
  gee_path          = "~/google_earth_engine_subsets/gee_subset/",
  data_path         = "~/data/gee_subsets/",
  method_interpol   = "linear",
  keep              = TRUE,
  overwrite_raw     = FALSE,
  overwrite_interpol= TRUE
  )

df_gee_modis_evi <- ingest(
  siteinfo= ingestr::siteinfo %>% 
    dplyr::filter(!(sitename %in% c("AU-GWW", "AU-Lox", "AU-Rob", "AU-TTE", "CN-Dan"))),
  source  = "gee",
  settings= settings_gee,
  verbose = FALSE
  ) 

df <- df_gee_modis_evi %>% 
  unnest(data) %>% 
  dplyr::select(sitename, date, evi = modisvar_interpol)

## quick check
df %>% 
  dplyr::filter(sitename == "AR-SLu") %>% 
  ggplot(aes(x = date, y = evi)) +
  geom_line()

write_csv(
  df, 
  path = "~/data/fluxnet_subsets/EVI_MOD13Q1_gee_subset.csv"
  )
```

Collect all plots.
```{r warning=FALSE, message=FALSE, eval = FALSE}
list_gg <- plot_fapar_ingestr(df_gee_modis_fpar, settings_gee)
#purrr::map(list_gg, ~print(.))
```


