---
title: "Description of how to get site data"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(tidyverse)
library(ingestr)
```


## Descriptions

This is an example for how to collect data to force the P-model (rsofun run) for a site where we don't have network measured data but instead extract climate forcing data from global files (WATCH-WFDEI and CRU).

## WATCH-WFDEI

Let's extract data for the location corresponding to FLUXNET site 'CH-Lae' (lon = 8.365, lat = 47.4781). This extracts from original WATCH-WFDEI files, provided as NetCDF (global, 0.5 degree resolution), provided as monthly files containing all days in each month. The data directory specified here (`dir = "~/data/watch_wfdei/"`) contains subdirectories with names containing the variable names (corresponding to the ones specified by the argument `getvars = list(temp = "Tair")`). We read data for just one year here (2007).
```{r message=FALSE, echo = T, results = 'hide'}
df_watch <- ingest_bysite(
  sitename  = "efg",  # can be any name
  source    = "watch_wfdei",
  getvars   = list(temp = "Tair", prec = "Rainf", vpd = "Qair", ppfd = "SWin"),
  dir       = "~/data/watch_wfdei/",
  timescale = "d",
  year_start = 2015,
  year_end  = 2015,
  lon       = 117.369258403,
  lat       = 30.026229678,
  verbose   = FALSE
  )
df_watch
```

## Local measured data

### Daily data

Read file with daily data and compare to WATCH-WFDEI.
```{r}
library(lubridate)
df_site <- read_csv("~/data/neecf/climate_data/site_ds_neecf_daily.csv") %>% 
  mutate(date = lubridate::ymd(paste0(as.character(year), "-01-01"))) %>% 
  mutate(date = date + months(month-1)) %>% 
  mutate(date = date + days(day-1)) %>% 
  rename(prec_obs = precipitation, temp_obs = temperature)

df_test <- df_watch %>% 
  left_join(df_site, by = "date")

df_test %>% 
  ggplot() +
  geom_line(aes(date, temp), color = "red") +
  geom_line(aes(date, temp_obs))

df_test %>% 
  ggplot() +
  geom_line(aes(date, prec), color = "red") +
  geom_line(aes(date, prec_obs))
```

Overwrite WATCH-WFDEI data if measured data is available.
```{r}
ddf <- df_test %>% 
  mutate(temp = ifelse(!is.na(temp_obs), temp_obs, temp),
         prec = ifelse(!is.na(prec_obs), prec_obs, prec))
```

Check.
```{r}
ddf %>% 
  ggplot() +
  geom_line(aes(date, temp), color = "red") +
  geom_line(aes(date, temp_obs))

ddf %>% 
  ggplot() +
  geom_line(aes(date, prec), color = "red") +
  geom_line(aes(date, prec_obs))
```


### Sub-daily data

Use sub-daily data to get daytime mean temperature and daytime mean VPD.
```{r}
hdf <- read_csv("~/data/neecf/climate_data/site_ds_neecf_halfhourly.csv") %>% 
  mutate(date_time = ymd_hm(date_time))

hdf %>% slice(1:100) %>% ggplot(aes(date_time, air_temp)) + geom_line()
```

Get daily mean and daytime mean air temperature
```{r}
ddf_hdf <- hdf %>% 
  mutate(date = date(date_time)) %>% 
  ungroup() %>% 
  group_by(date) %>% 
  summarise(temp_mean = mean(air_temp, na.rm = TRUE))
```
Compare.
```{r}
df_test <- ddf %>% 
  left_join(ddf_hdf, by = "date")

df_test %>% 
  ggplot() +
  geom_line(aes(date, temp_mean), color = "red") +
  geom_line(aes(date, temp))
```

### VPD

Assume relative humidity constant throughout the day, and use observed diurnal temperature variations to calculate diurnal VPD variations.
```{r}
hdf <- hdf %>% 
  mutate(date = ymd(date)) %>% 
  left_join(dplyr::select(ddf, date, qair), by = "date") %>% 
  rowwise() %>% 
  mutate(vpd = calc_vpd( qair = qair, tc = air_temp, tmin = NA, tmax = NA, patm = NA, elv = 330  ))

hdf %>% 
  slice(1:500) %>% 
  ggplot(aes(date_time, vpd)) +
  geom_line()

hdf %>% 
  ggplot(aes(x = vpd, y = ..density..)) +
  geom_histogram()
```

Get daily mean VPD based on hourly VPD.
```{r}
ddf_hdf <- hdf %>% 
  mutate(date = date(date_time)) %>% 
  ungroup() %>% 
  group_by(date) %>% 
  summarise(temp_mean = mean(air_temp, na.rm = TRUE),
            vpd_mean = mean(vpd, na.rm = TRUE))

ddf_midday_hdf <- hdf %>% 
  mutate(date = date(date_time)) %>% 
  dplyr::filter(hour(date_time) %in% 11:16) %>% 
  ungroup() %>% 
  group_by(date) %>% 
  summarise(temp_mean = mean(air_temp, na.rm = TRUE),
            vpd_mean = mean(vpd, na.rm = TRUE))
```

Check.
```{r}
ddf_hdf %>% ggplot(aes(date, vpd_mean)) + geom_line()

gg1 <- ggplot() +
  geom_line(data = hdf %>% dplyr::filter(month(date_time) %in% 6:8),
            aes(date_time, vpd))
gg2 <- ggplot() +
  geom_line(data = ddf_midday_hdf %>% dplyr::filter(month(date) %in% 6:8), 
            aes(date, vpd_mean))

gg3 <- ddf %>% 
  dplyr::filter(month(date) %in% 6:8) %>% 
  ggplot(aes(date, prec)) + 
  geom_line(color = "royalblue")

library(patchwork)
gg1 / gg2 / gg3
```

Complement daily data frame `ddf`.
```{r}
ddf <- ddf %>% 
  left_join(
    ddf_midday_hdf %>% 
      rename(vpd_midday = vpd_mean, temp_midday = temp_mean),
    by = "date"
  )
```

Check
```{r}
ddf %>% 
  ggplot() +
  geom_line(aes(date, temp_midday), color = "red") +
  geom_line(aes(date, temp))

ddf %>% 
  ggplot() +
  geom_line(aes(date, vpd_midday))

ddf %>% 
  ggplot() +
  geom_line(aes(date, ppfd))
```

## CRU TS

Cloud cover data.
```{r message=FALSE}
df_cru <- ingest_bysite(
  sitename  = "efg",  # can be any name
  source    = "cru",
  getvars   = list(ccov = "cld"),
  dir       = "~/data/cru/ts_4.01/",
  timescale = "d",
  year_start = 2015,
  year_end  = 2015,
  lon       = 117.369258403,
  lat       = 30.026229678,
  verbose   = FALSE
  )

df_cru
```

## CO2

Ingesting CO2 data is particularly simple. We can safely assume it's well mixed in the atmosphere (independent of site location), and we can use a annual mean value for all days in respective years, and use the same value for all
sites. Using the R package [climate](https://github.com/bczernecki/climate), we can load CO2 data from Mauna Loa directly into R. This is downloading data from [ftp://aftp.cmdl.noaa.gov/products/trends/co2/co2_mm_mlo.txt](ftp://aftp.cmdl.noaa.gov/products/trends/co2/co2_mm_mlo.txt). Here, `ingest()` is a wrapper for the function `climate::meteo_noaa_co2()`.
```{r message=FALSE}
df_co2 <- ingest_bysite(
  sitename  = "efg",
  source  = "co2_mlo",
  year_start= 2015,
  year_end  = 2015,
  verbose = FALSE
  )
```

## Combine

... and save.
```{r}
ddf <- ddf %>% 
  left_join(df_cru, by = "date") %>% 
  left_join(df_co2, by = "date")

write_csv(ddf, path = "~/data/neecf/climate_data/climate_data_collected.csv")
```