---
title: "ingest CRU"
author: "Koen Hufkens"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ingest CRU}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE, eval = FALSE}
library(ingestr)
library(dplyr)

# Below definition of analyse_modobs2() 
# was copy-pasted from: https://github.com/stineb/rbeni/blob/2f9d26d0a286c550cb90ee9d30a1f2b6c3b112f6/R/analyse_modobs2.R

#' Analyse modelled values versus observed data.
#'
#' Calculates a set of performance statistics and optionally creates plots of modelled
#' versus observed values.
#'
#' @param df A data frame containing columns with names corresponding to arguments
#' \code{mod} and \code{obs}
#' @param mod A character string specifying the variable name (column) of the
#' modelled (simulated) values in data frame \code{df}.
#' @param obs A character string specifying the variable name (column) of the
#' observed values in data frame \code{df}.
#' @param type If \code{"points"}, uses \code{geom_points()}, if \code{"hex"}
#' uses \code{ggplot2::geom_hex()}, if \code{"heat"} uses adjusted
#' \code{geom_points()} with color indicating density, if \code{"density"} uses
#' \code{stat_density_2d()} to draws polygos of equal density.
#' @param filnam A character string specifying the name of the file containing
#' the plot. Defaults to \code{NA} (no file is created).
#' @param relative A logical specifying whether the relative RMSE and bias (after
#' division by the mean) is to be showed in the subtitle labels.
#' @param shortsubtitle A boolean specifying whether to display a reduced set of metrics
#' in the subtitle.
#' @param rsquared A boolean specifying whether to display R-squared and the RMSE
#' (if \code{TRUE}) or the r (Pearson's correlation coefficient) and the p (p-value of
#' test of significance of correlation, if \code{TRUE}). Defaluts to \code{TRUE}.
#' @param plot_subtitle A boolean specifying whether to display any metrics. Defaults
#' to \code{TRUE}.
#' @param plot_linmod A boolean specifying whether to display the fitted linear
#' regression as a red line. Defaults to \code{TRUE}.
#' @param plot_legend A boolean specifying whether to display a legend for the colors.
#' Defaults to \code{TRUE} if \code{type} is one of  \code{"heat"},  \code{"hex"}, or
#' \code{"density"}.
#' @param label A boolean specifying whether points should be labelled using ggrepel.
#' Defaults to \code{FALSE}. Only available for \code{type == "points"}. Use argument
#' \code{nlabels} to specify how many points should be labelled, starting with points
#' that have the largest residuals from the linear regression fit.
#' @param id A character string specifying the column name that identifies the points.
#' The column's values must be of type integer and is used to label points in case of
#' \code{label = TRUE}.
#' @param nlabels An integer specifying how many points to be labelled, starting with points
#' that have the largest residuals from the linear regression fit. Only available
#' for \code{type == "points"}. Defaults to one.
#'
#' @export
#'
#' @examples
#'
analyse_modobs2 <- function(
  df,
  mod,
  obs,
  type       = "points",
  filnam     = NA,
  relative   = FALSE,
  xlim       = NULL,
  ylim       = NULL,
  use_factor = NULL,
  shortsubtitle = FALSE,
  rsquared    = TRUE,
  plot_subtitle = TRUE,
  plot_linmod = TRUE,
  plot_legend = TRUE,
  label       = FALSE,
  id          = NULL,
  nlabels     = 1,
  ...
  ){

  require(ggplot2)
  require(dplyr)
  require(LSD)
  require(ggthemes)
  require(RColorBrewer)

  #if (identical(filnam, NA)) filnam <- "analyse_modobs.pdf"

  ## rename to 'mod' and 'obs' and remove rows with NA in mod or obs
  if (label){
    df <- df %>%
      as_tibble() %>%
      ungroup() %>%
      dplyr::select(mod=mod, obs=obs, id=!!id) %>%
      tidyr::drop_na(mod, obs)

  } else {
    df <- df %>%
      as_tibble() %>%
      ungroup() %>%
      dplyr::select(mod=mod, obs=obs) %>%
      tidyr::drop_na(mod, obs)

  }

  ## get linear regression (coefficients)
  linmod <- lm( obs ~ mod, data=df )

  ## construct metrics table using the 'yardstick' library
  df_metrics <- df %>%
    yardstick::metrics(obs, mod) %>%
    dplyr::bind_rows( tibble( .metric = "n",        .estimator = "standard", .estimate = summarise(df, numb=n()) %>% unlist() ) ) %>%
    dplyr::bind_rows( tibble( .metric = "slope",    .estimator = "standard", .estimate = coef(linmod)[2]) ) %>%
    # dplyr::bind_rows( tibble( .metric = "nse",      .estimator = "standard", .estimate = hydroGOF::NSE( obs, mod, na.rm=TRUE ) ) ) %>%
    dplyr::bind_rows( tibble( .metric = "mean_obs", .estimator = "standard", .estimate = summarise(df, mean=mean(obs, na.rm=TRUE)) %>% unlist() ) ) %>%
    dplyr::bind_rows( tibble( .metric = "prmse",    .estimator = "standard",
                       .estimate = dplyr::filter(., .metric=="rmse") %>% dplyr::select(.estimate) %>% unlist() /
                         dplyr::filter(., .metric=="mean_obs") %>% dplyr::select(.estimate) %>% unlist() ) ) %>%
    dplyr::bind_rows( tibble( .metric = "pmae",    .estimator = "standard",
                       .estimate = dplyr::filter(., .metric=="mae") %>% dplyr::select(.estimate) %>% unlist() /
                         dplyr::filter(., .metric=="mean_obs") %>% dplyr::select(.estimate) %>% unlist() ) ) %>%
    dplyr::bind_rows( tibble( .metric = "bias",  .estimator = "standard", .estimate = dplyr::summarise(df, mean((mod-obs), na.rm=TRUE    )) %>% unlist() ) ) %>%
    dplyr::bind_rows( tibble( .metric = "pbias", .estimator = "standard", .estimate = dplyr::summarise(df, mean((mod-obs)/obs, na.rm=TRUE)) %>% unlist() ) ) %>%
    dplyr::bind_rows( tibble( .metric = "cor",   .estimator = "standard", .estimate = cor(df$mod, df$obs, method = "pearson") ) ) %>%
    dplyr::bind_rows( tibble( .metric = "cor_p", .estimator = "standard", .estimate = cor.test(df$mod, df$obs, method = "pearson")$p.value ) )

  rsq_val <- df_metrics %>% dplyr::filter(.metric=="rsq") %>% dplyr::select(.estimate) %>% unlist() %>% unname()
  rmse_val <- df_metrics %>% dplyr::filter(.metric=="rmse") %>% dplyr::select(.estimate) %>% unlist() %>% unname()
  mae_val <- df_metrics %>% dplyr::filter(.metric=="mae") %>% dplyr::select(.estimate) %>% unlist() %>% unname()
  bias_val <- df_metrics %>% dplyr::filter(.metric=="bias") %>% dplyr::select(.estimate) %>% unlist() %>% unname()
  slope_val <- df_metrics %>% dplyr::filter(.metric=="slope") %>% dplyr::select(.estimate) %>% unlist() %>% unname()
  n_val <- df_metrics %>% dplyr::filter(.metric=="n") %>% dplyr::select(.estimate) %>% unlist() %>% unname()
  cor_val <- df_metrics %>% dplyr::filter(.metric=="cor") %>% dplyr::select(.estimate) %>% unlist() %>% unname()
  cor_p_val <- df_metrics %>% dplyr::filter(.metric=="cor_p") %>% dplyr::select(.estimate) %>% unlist() %>% unname()

  if (relative){
    rmse_val <- rmse_val / mean(df$obs, na.rm = TRUE)
    bias_val <- bias_val / mean(df$obs, na.rm = TRUE)
  }

  rsq_lab <- format( rsq_val, digits = 2 )
  rmse_lab <- format( rmse_val, digits = 3 )
  mae_lab <- format( mae_val, digits = 3 )
  bias_lab <- format( bias_val, digits = 3 )
  slope_lab <- format( slope_val, digits = 3 )
  n_lab <- format( n_val, digits = 3 )
  cor_lab <- format( cor_val, digits = 3 )
  cor_p_lab <- format( cor_p_val, digits = 3 )

  results <- tibble( rsq = rsq_val, rmse = rmse_val, mae = mae_val, bias = bias_val, slope = slope_val, n = n_val )

  if (shortsubtitle){
    if (rsquared){
      subtitle <- bquote(
        italic(R)^2 == .(rsq_lab) ~~
        RMSE == .(rmse_lab)
      )
    } else {
      subtitle <- bquote(
        italic(r) == .(cor_lab) ~~
        italic(p) == .(cor_p_lab)
      )
    }
  } else {
    subtitle <- bquote( italic(R)^2 == .(rsq_lab) ~~
                          RMSE == .(rmse_lab) ~~
                          bias == .(bias_lab) ~~
                          slope == .(slope_lab) ~~
                          italic(N) == .(n_lab) )
  }

  if (type=="heat"){

    # if (!identical(filnam, NA)) dev.off()
    # source("~/LSD/R/LSD.heatscatter.R")

    gg <- heatscatter(
                  df$mod,
                  df$obs,
                  xlim=xlim,
                  ylim=ylim,
                  main="",
                  ggplot=TRUE )

    gg <- gg +
      geom_abline(intercept=0, slope=1, linetype="dotted") +
      theme_classic() +
      labs(x = mod, y = obs)

    if (plot_linmod) gg <- gg + geom_smooth(method='lm', color="red", size=0.5, se=FALSE)
    if (plot_subtitle) gg <- gg + labs(subtitle = subtitle)
    if (!plot_legend) gg <- gg + theme(legend.position = "none")

    if (!identical(filnam, NA)) {
      ggsave(filnam, width=5, height=5)
    }

  } else if (type=="hex"){

    ## ggplot hexbin
    gg <- df %>%
      ggplot2::ggplot(aes(x=mod, y=obs)) +
      geom_hex(bins = 100) +
      scale_fill_gradientn(
        colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5),
        trans = "log") +
      geom_abline(intercept=0, slope=1, linetype="dotted") +
      # coord_fixed() +
      # xlim(0,NA) +
      # ylim(0,NA) +
      theme_classic() +
      labs(x = mod, y = obs)

    if (plot_subtitle) gg <- gg + labs(subtitle = subtitle)
    if (plot_linmod) gg <- gg + geom_smooth(method='lm', color="red", size=0.5, se=FALSE)
    if (!plot_legend) gg <- gg + theme(legend.position = "none")

    if (!identical(filnam, NA)) {
      ggsave(filnam, width=5, height=5)
    }

  } else if (type=="points"){

    if (label){
      df <- df %>%
        dplyr::mutate(.res = mod - obs) %>%
        dplyr::mutate(.absres = abs(.res)) %>%
        dplyr::arrange(desc(.absres)) %>%
        dplyr::mutate(.rankres = 1:n()) %>%
        dplyr::mutate(.dolab = ifelse(.rankres <= nlabels, TRUE, FALSE))

      ## points with labels
      library(ggrepel)
      gg <- df %>%
        ggplot(aes(x=mod, y=obs, label = ifelse(.dolab, id, ""))) +
        geom_point() +
        geom_label_repel(min.segment.length = 0, seed = 42, box.padding = 0.5) +
        geom_point(data = dplyr::filter(df, .dolab), color = "red") +
        geom_abline(intercept=0, slope=1, linetype="dotted") +
        theme_classic() +
        labs(x = mod, y = obs)

    } else {
      ## points
      gg <- df %>%
        ggplot(aes(x=mod, y=obs)) +
        geom_point() +
        geom_abline(intercept=0, slope=1, linetype="dotted") +
        theme_classic() +
        labs(x = mod, y = obs)

    }


    if (plot_subtitle) gg <- gg + labs(subtitle = subtitle)
    if (plot_linmod) gg <- gg + geom_smooth(method='lm', color="red", size=0.5, se=FALSE)

    if (!identical(filnam, NA)) {
      ggsave(filnam, width=5, height=5)
    }

  } else if (type=="density"){

    ## density as raster
    gg <- df %>%
      ggplot(aes(x=mod, y=obs)) +

      stat_density_2d(
        geom = "raster", #the geometric object to display the data (in this case: rectangles)
        aes(fill = after_stat(density)), #using `density`, a variable calculated by the stat
        contour = FALSE
      ) +

      scale_fill_gradientn(colours = colorRampPalette( c("white", "gray65", "navy", "red", "yellow"))(6),
                           guide = FALSE) +

      geom_abline(intercept=0, slope=1, linetype="dotted") +
      # coord_fixed() +
      # xlim(0,NA) +
      # ylim(0,NA) +
      theme_classic() +
      labs(x = mod, y = obs)

    if (plot_subtitle) gg <- gg + labs(subtitle = subtitle)
    if (plot_linmod) gg <- gg + geom_smooth(method='lm', color="red", size=0.5, se=FALSE)
    if (!plot_legend) gg <- gg + theme(legend.position = "none")

    if (!identical(filnam, NA)) {
      ggsave(filnam, width=5, height=5)
    }

  }

  return(list(df_metrics=df_metrics, gg=gg, linmod=linmod, results = results))
}
```

CRU TS provides monthly climate fields at 0.5 degree resolution from 1901 to today. Frequent updates are made available. CRU TS 4.05 contains data up to 2018. ingestr converts CRU variables to rsofun standard variables and units that can then be used as forcing for rsofun. The following describes what precursor variables are used for each rsofun standard variable (and how). 

| rsofun standard variable   | (Precursor) CRU variable name(s)  | Remark
|---                         |---                                |---
| tmin                       | tmn                               |
| tmax                       | tmx                               |
| prec                       | prc, wtd                          | Weather generator conserving monthly sums and number of monthly wet days
| vpd                        | vap, tmin, tmax                   | Using `calc_vpd()` 
| ccov                       | cld                               | 


```{r warning=FALSE, eval = FALSE}
## get monthly data (no temporal downscaling - original extracted values)
mdf <- ingest_bysite(
  sitename  = "CH-Lae",
  source    = "cru",
  getvars   = c("tmax", "tmin", "prec", "vpd"),
  # dir       = "/data/archive/cru_NA_2021/data/",
  dir       = "/data/archive/cru_harris_2024/data/",
  timescale = "m",
  year_start = 1901,
  year_end  = 2018,
  lon       = 8.365,
  lat       = 47.4781,
  verbose   = FALSE
  )

## get daily data (with temporal downscaling)
ddf <- ingest_bysite(
  sitename  = "CH-Lae",
  source    = "cru",
  getvars   = c("tmax", "tmin", "prec", "vpd"),
  dir       = "/data/archive/cru_harris_2024/data/",
  timescale = "d",
  year_start = 1901,
  year_end  = 2018,
  lon       = 8.365,
  lat       = 47.4781,
  verbose   = FALSE
  )
```

## Check temporal downscaling

The temporal downscaling conserves monthly means. The following shows monthly tmin values aggregated from downscaled daily values versus values directly extracted from the original files.
```{r warning=FALSE, eval = FALSE}
mdf_test <- ddf %>% 
  mutate(year = lubridate::year(date), month = lubridate::month(date)) %>% 
  group_by(year, month) %>% 
  summarise(tmin = mean(tmin)) %>% 
  rename(tmin_agg = tmin) %>% 
  ungroup() %>% 
  left_join(mdf %>% 
              select(year, month, tmin_orig = tmin))

gg <- mdf_test %>% analyse_modobs2("tmin_orig", "tmin_agg")
gg$gg + labs(x = "Original monthly tmin (deg C)", y = "Aggregated monthly tmin (deg C)")
```

For precipitation, the temporal downscaling conserves monthly totals and distributes precipitation to the given number of wet days (also provided by CRU as the number of wet days per month).
```{r warning=FALSE, eval = FALSE}
mdf_test <- ddf %>% 
  mutate(year = lubridate::year(date), month = lubridate::month(date)) %>% 
  group_by(year, month) %>% 
  summarise(prec = mean(prec)) %>% 
  rename(prec_agg = prec) %>% 
  ungroup() %>% 
  left_join(mdf %>% 
              select(year, month, prec_orig = prec))

gg <- mdf_test %>% analyse_modobs2("prec_orig", "prec_agg")
gg$gg + labs(x = "Original monthly prec (mm)", y = "Aggregated monthly prec (mm)")
```

Monthly means are not conserved for VPD. This is because CRU TS provides vapour pressure (VAP) data and VPD is calculated by ingestr as 
$$
VPD = (f(VAP, TMIN) + f(VAP, TMAX))/2
$$

Where $f$ is a non-linear function (`calc_vpd()`) and VAP, TMIN, and TMAX are either monthly mean values in case of `timescale = "m"` or daily values (conserved monthly means) in case of `timescale = "d"`,

```{r warning=FALSE, eval = FALSE}
mdf_test <- ddf %>% 
  mutate(year = lubridate::year(date), month = lubridate::month(date)) %>% 
  group_by(year, month) %>% 
  summarise(vpd = mean(vpd)) %>% 
  rename(vpd_agg = vpd) %>% 
  ungroup() %>% 
  left_join(mdf %>% 
              select(year, month, vpd_orig = vpd))

gg <- mdf_test %>% analyse_modobs2("vpd_orig", "vpd_agg")
gg$gg + labs(x = "Original monthly VPD (Pa)", y = "Aggregated monthly VPD (Pa)")
```

## Bias correction with WorldClim

Bias correction based on high-resolution WorldClim 1970-2000 monthly climatology is available for variables temp, prec, and vpd.
```{r warning=FALSE, eval = FALSE}
## get monthly data (no temporal downscaling - original extracted values)
mdf_corr <- ingest_bysite(
  sitename  = "CH-Lae",
  source    = "cru",
  getvars   = c("temp", "tmin", "tmax", "prec", "vpd", "ccov"),
  dir       = "/data/archive/cru_harris_2024/data/",
  timescale = "m",
  year_start = 1901,
  year_end  = 2018,
  lon       = 8.365,
  lat       = 47.4781,
  verbose   = FALSE,
  settings  = list(correct_bias = "worldclim", dir_bias = "/data/archive/worldclim_fick_2017/data")
  )

## get daily data (with temporal downscaling)
ddf_corr <- ingest_bysite(
  sitename  = "CH-Lae",
  source    = "cru",
  getvars   = c("temp", "tmin", "tmax", "prec", "vpd", "ccov"),
  dir       = "/data/archive/cru_harris_2024/data/",
  timescale = "d",
  year_start = 1901,
  year_end  = 2018,
  lon       = 8.365,
  lat       = 47.4781,
  verbose   = FALSE,
  settings  = list(correct_bias = "worldclim", dir_bias = "/data/archive/worldclim_fick_2017/data")
  )
```
Check conservation of precipitation means after bias correction.
```{r warning=FALSE, eval = FALSE}
mdf_test <- ddf_corr %>% 
  mutate(year = lubridate::year(date), month = lubridate::month(date)) %>% 
  group_by(year, month) %>% 
  summarise(prec = mean(prec)) %>% 
  rename(prec_agg = prec) %>% 
  ungroup() %>% 
  left_join(mdf_corr %>% 
              select(year, month, prec_orig = prec))

gg <- mdf_test %>% analyse_modobs2("prec_orig", "prec_agg")
gg$gg + labs(x = "Original monthly prec (deg C)", y = "Aggregated monthly prec (deg C)")
```

Check conservation of VPD means after bias correction.
```{r warning=FALSE, eval = FALSE}
mdf_test <- ddf_corr %>% 
  mutate(year = lubridate::year(date), month = lubridate::month(date)) %>% 
  group_by(year, month) %>% 
  summarise(vpd = mean(vpd)) %>% 
  rename(vpd_agg = vpd) %>% 
  ungroup() %>% 
  left_join(mdf_corr %>% 
              select(year, month, vpd_orig = vpd))

gg <- mdf_test %>% analyse_modobs2("vpd_orig", "vpd_agg")
gg$gg + labs(x = "Original monthly vpd (deg C)", y = "Aggregated monthly vpd (deg C)")
```

## Check against station data

Comparison of bias-corrected data to FLUXNET site-level observations. For CH-Lae, this is available for 2004-2014. Visualize for three years (2012-2014). 
Get FLUXNET data.
```{r warning=FALSE, eval = FALSE}
ddf_fluxnet <- ingestr::ingest(
  siteinfo  = siteinfo_fluxnet2015 %>% dplyr::filter(sitename == "CH-Lae"),
  source    = "fluxnet",
  getvars   = list(temp = "TA_F_DAY", prec = "P_F", vpd  = "VPD_F_DAY", ppfd = "SW_IN_F", patm = "PA_F"),
  dir       = "/data/scratch/bstocker/FLUXNET-2015_Tier1/20191024/DD/",
  settings  = list(dir_hh = "/data/scratch/bstocker/FLUXNET-2015_Tier1/20191024/HH/", getswc = FALSE),
  timescale = "d"
  ) %>% 
  tidyr::unnest(data)
```

Looks fine for temperature.
```{r warning=FALSE, eval = FALSE}
ggplot() +
  geom_line(data = ddf_fluxnet %>% 
              dplyr::filter(lubridate::year(date) %in% 2012:2014), 
            aes(date, temp)) +
  geom_line(data = ddf_corr %>% 
              dplyr::filter(lubridate::year(date) %in% 2012:2014), 
            aes(date, temp),
            color = "red")

out <- ddf_fluxnet %>% 
  mutate(year = lubridate::year(date), month = lubridate::month(date)) %>% 
  group_by(year, month) %>% 
  summarise(temp_fluxnet = mean(temp)) %>% 
  left_join(ddf_corr %>% 
              mutate(year = lubridate::year(date), month = lubridate::month(date)) %>% 
              group_by(year, month) %>% 
              summarise(temp_cru_wc = mean(temp)),
            by = c("year", "month")) %>% 
  analyse_modobs2("temp_fluxnet", "temp_cru_wc")
out$gg
```

Looks fine for precipitation. Compare monthly means - not bad at all!
```{r warning=FALSE, eval = FALSE}
out <- ddf_fluxnet %>% 
  mutate(year = lubridate::year(date), month = lubridate::month(date)) %>% 
  group_by(year, month) %>% 
  summarise(prec_fluxnet = mean(prec)) %>% 
  left_join(ddf_corr %>% 
              mutate(year = lubridate::year(date), month = lubridate::month(date)) %>% 
              group_by(year, month) %>% 
              summarise(prec_cru_wc = mean(prec)),
            by = c("year", "month")) %>% 
  analyse_modobs2("prec_fluxnet", "prec_cru_wc")
out$gg
```

Looks fine for VPD
```{r warning=FALSE, eval = FALSE}
ggplot() +
  geom_line(data = ddf_fluxnet %>% 
              dplyr::filter(lubridate::year(date) %in% 2012:2014), 
            aes(date, vpd)) +
  geom_line(data = ddf_corr %>% 
              dplyr::filter(lubridate::year(date) %in% 2012:2014), 
            aes(date, vpd),
            color = "red")

out <- ddf_fluxnet %>% 
  mutate(year = lubridate::year(date), month = lubridate::month(date)) %>% 
  group_by(year, month) %>% 
  summarise(vpd_fluxnet = mean(vpd)) %>% 
  left_join(ddf_corr %>% 
              mutate(year = lubridate::year(date), month = lubridate::month(date)) %>% 
              group_by(year, month) %>% 
              summarise(vpd_cru_wc = mean(vpd)),
            by = c("year", "month")) %>% 
  analyse_modobs2("vpd_fluxnet", "vpd_cru_wc")
out$gg
```