# Aim

In this vignette, I try to extract the T air and Q air, to obtain patm, vpd and ppfd

Fapar still to do

I modified only the function extract_pointdata_allsites_shp and ingest_globalfields_watch_byvar. The latter is modified in such a way that (in theory) using the coordinates points instead of the shapefile doesn't change anything. The first function is called only when there is a shapefile.

Now we'll have a look at the new function extract_pointdata_allsites_shp

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(terra)
library(lubridate)
library(ingestr)
# data preparation, currently only on my folder

# will take at random 10 catchements and see what happens

vector_path <- "~/data_scratch/shapefiles/camels/camels_basin_shapes.shp"

shapefile <- terra::vect(vector_path)

shapefile <- shapefile[300:309,]

siteinfo <- read.csv("~/data_scratch/camels_timeseries/camels/attributes_other_camels.csv")

siteinfo <- siteinfo[siteinfo$gauge_id %in% shapefile$gauge_id,]

# we add the year of interest, for example 2004

date_start = lubridate::ymd(paste0(2005, "-01-01"))
date_end = lubridate::ymd(paste0(2005, "-12-31"))


siteinfo <- siteinfo |>
  dplyr::mutate(date_start = date_start) |>
  dplyr::mutate(date_end = date_end) |>
  mutate(sitename = gauge_id) |>
  rename(lon = gauge_lon,
         lat = gauge_lat)

```


## CO2

Get CO2 before so I can put directly in driver_data

```{r}
df_co2 <- ingest_bysite(
  sitename  = "anyone",
  source  = "co2_mlo",
  year_start= lubridate::year(date_start),
  year_end  = lubridate::year(date_end),
  verbose = FALSE
  )
```



```{r}
# laod CSV data and driver data preparation


timeseries <- list.files(path = "~/data_scratch/camels_timeseries/camels/csv/", pattern = "\\.csv$", full.names = T)


timeseries <- timeseries[substr(timeseries,60,74) %in% siteinfo$gauge_id]

driver_data <- NULL

for(i in timeseries){
  sitename <- substr(i,60,74)
  
  params_siml <- rsofun::p_model_drivers$params_siml[[1]] |>
    mutate(use_gs = TRUE, use_phydro = FALSE, use_pml= TRUE) |>
    nest(params_siml = c(spinup, spinupyears, recycle, outdt, ltre,  ltne,  ltrd,  ltnd,
                         lgr3,  lgn3,  lgr4 ,
                         use_gs, use_phydro, use_pml))
  
  site_info <- siteinfo[siteinfo$gauge_id == sitename,] |>
    select(lat,lon) |>
    mutate(canopy_height =12, # TODO 
           reference_height = 10) |> # I select 10 beacuse the wind velocity is measured at 10 m
    nest(site_info = c(lat,lon,canopy_height,reference_height))
  
  forcing <- read_csv(i)
  forcing <- forcing |>
    filter(lubridate::date(date) >= date_start, lubridate::date(date) <= date_end) |> 
    rename(rain = total_precipitation_sum,
           runoff = streamflow,
           tmin = temperature_2m_min,
           tmax = temperature_2m_max,
           temp = temperature_2m_mean,
           vwind = u_component_of_wind_10m_max) |>
    mutate(netrad = NA,
           snow = 0,
           co2 = df_co2$co2,
           fapar = 1, # TODO
           rain = rain / (24*60*60)) |>
    select(date,temp,netrad,snow,rain,tmin,tmax,vwind,fapar,co2,runoff) |>
    filter(!(lubridate::month(date) == 2 & lubridate::day(date) == 29)) |>
    nest(forcing = c(date,temp,netrad,snow,rain,tmin,tmax,vwind,fapar,co2,runoff))
  
  tmp <- data.frame(sitename = sitename,
                    params_siml = tibble(params_siml),
                    site_info = tibble(site_info),
                    forcing = tibble(forcing))
  
  driver_data <- rbind(driver_data,tmp)
  
}

```


```{r}
df_out <- tibble()

getvars <- c("vpd",'ppfd')

source <- "watch_wfdei"

dir <- "/data/archive/wfdei_weedon_2014/data/"

timescale <- "d"

df_out <- ingest_globalfields(siteinfo = siteinfo,source = source,getvars = getvars,
                               dir = dir, timescale= timescale,is_shapefile = T, shapefile = shapefile)


driver_data <- 
  left_join(driver_data |>
  unnest(forcing),
  df_out |>
    dplyr::select(sitename,date,ppfd,patm,vpd),
  by =c("sitename","date")) |>
  nest(forcing = c(date,temp,vpd,ppfd,netrad,patm,snow,rain,tmin,tmax,vwind,fapar,co2,runoff))

```


## extract WHC

To further prove that the shapefile works, I extract the WHC using the coordinates and the shapefile,
They should be different

```{r}
df_out <- tibble()

getvars <- NULL #c("vpd",'ppfd')

source <- "stocker23"

dir <- "/data/archive/whc_stocker_2023/data/"


# WHC point scale
df_out <- ingest_globalfields(siteinfo = siteinfo,source = source,getvars = getvars,
                               dir = dir, timescale= timescale,is_shapefile = F, shapefile = NULL)

print(df_out)

```

```{r}
# based on shapefile

df_out <- tibble()

df_out <- ingest_globalfields(siteinfo = siteinfo,source = source,getvars = getvars,
                               dir = dir, timescale= timescale,is_shapefile = T, shapefile = shapefile)

driver_data <- 
  left_join(driver_data |>
  unnest(site_info),
  df_out ,
  by ="sitename")|>
  nest(site_info = c(lat,lon,whc,canopy_height,reference_height))

print(df_out)
```

## Elevation

```{r}
df_out <- tibble()

source <- "etopo1"

dir <- "/data/archive/etopo_NA_NA/data/"

df_out <- ingest_globalfields(siteinfo = siteinfo,source = source,getvars = getvars,
                               dir = dir, timescale= timescale,is_shapefile = T, shapefile = shapefile)


driver_data <- 
  left_join(driver_data |>
  unnest(site_info),
  df_out ,
  by ="sitename")|>
  nest(site_info = c(lat,lon,elv,whc,canopy_height,reference_height))

```


## CLoud coverage (to fix)

```{r}
df_cld <- ingestr:::extract_pointdata_allsites_shp(
  filename = "/data/archive/cru_harris_2024/data/cru_ts4.08.1901.2023.cld.dat.nc",
  df_shapefile = shapefile,
  get_time = T,
  year_arg = NA_integer_, month_arg = NA_integer_ # only used for WFDEI in combination with get_time
) 

df_cld <- df_cld |> unnest(data) |>
  filter(varnam == "cld") |>
  filter(date > date_start & date < date_end) |>
  mutate(value = value /100) |>
  mutate(date = floor_date(date, "month"))


df_cld <- df_cld |>
  group_by(sitename) |>
  mutate(date = purrr::map(date, ~ seq(.x, ceiling_date(.x, "month") - days(1), by = "day"))) |>
  unnest(date) |>
  unnest(sitename) |>
  select(sitename,date, value) |>
  rename(ccov = value)

driver_data <- left_join(
  driver_data |> 
    unnest(forcing),
  df_cld, by=c("sitename","date")
  ) |> 
  nest(forcing =c(date,temp,vpd,ppfd,netrad,patm,snow,rain,tmin,tmax,vwind,fapar,co2,ccov,runoff))
```



```{r}
params_modl <- list(
    kphio              = 0.04998,    # setup ORG in Stocker et al. 2020 GMD
    kphio_par_a        = 0.0,        # set to zero to disable temperature-dependence of kphio
    kphio_par_b        = 1.0,
    soilm_thetastar    = 0.6 * 240,  # to recover old setup with soil moisture stress
    beta_unitcostratio = 146.0,
    rd_to_vcmax        = 0.014,      # value from Atkin et al. 2015 for C3 herbaceous
    tau_acclim         = 30.0,
    kc_jmax            = 0.41,
    gw_calib =2
  )

# run the model for these parameters
output <- rsofun::runread_pmodel_f(
   driver_data,
  par = params_modl
  )
```



